---
title: "Untitled"
author: "zexu chen"
date: "2018/03/01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(simmer)
library(dplyr)
library(magrittr)
```

arrival Rate= 9.451 patients/hour

Average Triage Time= 15.353 mins
StDev=17.625

Average RLOS= 172.684 mins
StDev= 149.819

Upstairs Time (Eval complete to time left ED): 87.836 mins
StDev = 160.521 

Percentage of arrivals:
Discharged: 77.69%
Admits: 18.18%
LAMA: 0.04%
LWBS: 1.90%
Not specified: 2.18%


patient = #arrival distribution?
room = 42
indocator = #labs, radiology, pharmmacy, consults, discharged/admitted

```{r}

numNurse = 2
numRoom = 42
numReplications = 30

```

#first version

patient <- trajectory("patient's path ") %>%
  seize("triage nurse", amount=1) %>% 
  timeout(function() rexp(n = 1, rate = 0.085))%>% 
  release("triage nurse", amount = 1)%>%
  seize("room", amount = 1)%>%
  timeout(function() rgamma(n = 1, shape = 2.618 ,rate = 0.016 ))%>%
  release("room",1)

 


#2 types of patients: admitted and discharged
#each get one trajectory
```{r}
patient1 <- trajectory("admitted ") %>%
  seize("triage nurse", amount=1) %>% 
  timeout(function() rexp(n = 1, rate = 0.085))%>% 
  release("triage nurse", amount = 1)%>%
  seize("room", amount = 1)%>%
  timeout(function() rgamma(n = 1, shape = 5.68 ,rate = 0.021))%>%
  release("room",1)


patient2 <- trajectory("discharged ") %>%
  seize("triage nurse", amount=1) %>% 
  timeout(function() rexp(n = 1, rate = 0.085))%>% 
  release("triage nurse", amount = 1)%>%
  seize("room", amount = 1)%>%
  timeout(function() rgamma(n = 1, shape = 2.82 ,rate = 0.02 ))%>%
  release("room",1)
```



#with branch
50-50 chance
patient <- trajectory("patient's path ") %>%
        branch(option = function() sample(1:2, 1), continue = c(T, F), 
        trajectory("discharged")%>%
          seize("triage nurse", amount=1) %>% 
          timeout(function() rexp(n = 1, rate = 0.085))%>% 
          release("triage nurse", amount = 1)%>%
          seize("room", amount = 1)%>%
          timeout(function() rgamma(n = 1, shape = 2.82 ,rate = 0.02 ))%>%
          release("room",amount = 1),
        trajectory("admitted")%>%
          seize("triage nurse", amount=1) %>% 
          timeout(function() rexp(n = 1, rate = 0.085))%>% 
          release("triage nurse", amount = 1)%>%
          seize("room", amount = 1)%>%
          timeout(function() rgamma(n = 1, shape = 5.68 ,rate = 0.021))%>%
          release("room",1)
          )


patient <- trajectory("patient's path ") %>%
  seize("triage nurse", amount=1) %>% 
  timeout(function() rexp(n = 1, rate = 0.085))%>% 
  release("triage nurse", amount = 1)%>%
  seize("room", amount = 1)%>%
  timeout(function() rgamma(n = 1, shape = 2.618 ,rate = 0.016 ))%>%
  release("room",1)



#with ESIs 
data needed :
shape and rate for the case admitted and discharged
arrival rates classified by ESIs and by time period
 


```{r}
hospital<-
  simmer("hospital") %>%
  add_resource("triage nurse",numNurse) %>%
  add_resource("room", numRoom)%>%
  #schedule(c(6, 12, 18, 24), c(3, 5, 2, 1), period=24)
  
  
  
  add_generator("patientESI1-a", patient1, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI1-b", patient1, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI1-c", patient1, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI1-d", patient1, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI2-a", patient1, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI2-b", patient1, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI2-c", patient1, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI2-d", patient1, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI3-a", patient1, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI3-b", patient1, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI3-c", patient1, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI3-d", patient1, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI4-a", patient1, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI4-b", patient1, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI4-c", patient1, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI4-d", patient1, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI5-a", patient1, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI5-b", patient1, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI5-c", patient1, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI5-d", patient1, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI0-a", patient1, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI0-b", patient1, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI0-c", patient1, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI0-d", patient1, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI1-e", patient2, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI1-f", patient2, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI1-g", patient2, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI1-h", patient2, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI2-e", patient2, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI2-f", patient2, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI2-g", patient2, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI2-h", patient2, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI3-e", patient2, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI3-f", patient2, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI3-g", patient2, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI3-h", patient2, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI4-e", patient2, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI4-f", patient2, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI4-g", patient2, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI4-h", patient2, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI5-e", patient2, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI5-f", patient2, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI5-g", patient2, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI5-h", patient2, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI0-e", patient2, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI0-f", patient2, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI0-g", patient2, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI0-h", patient2, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%

#generate the arrivals according to time and ESIs
#can using from_to to change the parameters

 run(until = 1440)

```


# initial version, 4 time periods
hospital<-
  simmer("hospital") %>%
  add_resource("triage nurse",numNurse) %>%
  add_resource("room", numRoom)%>%
  #schedule(c(6, 12, 18, 24), c(3, 5, 2, 1), period=24)
  
  
  
  add_generator("patienta", patient, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientb", patient, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientc", patient, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientd", patient, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  
  run(until = 1440)




There are three types of plot implemented with different metrics available:

a.
Plot of resources. Two metrics available:
the usage of a resource over the simulation time frame.
the utilization of specified resources in the simulation.
b.
Plot of arrivals. Three metrics available:
activity time.
waiting time.
flow time.
c.
Plot of attributes.


```{r}

#install.packages("simmer.plot")
```



#resource(triage nurse, room) utilization

The top of the bar shows the median utilization.

```{r}
library(simmer.plot)

resources <- get_mon_resources(hospital)
plot(resources, metric = "utilization")

```


#time uitilization(nurse & room, intervals, not timestamp)


```{r}
arrivals <- get_mon_arrivals(hospital, per_resource = T)
table(arrivals$resource)
# Let's see the distributions
ggplot(arrivals) + geom_histogram(aes(x=activity_time)) + facet_wrap(~resource)
 
``` 



```{r}
hospital %>%
  get_mon_arrivals%>%
  dplyr::mutate(service_start_time = end_time - activity_time) %>%
  mutate(wait_time = service_start_time - start_time)%>%
  dplyr::arrange(start_time)
  
```

```{r}
hospital%>%
  get_mon_resources %>%
  dplyr::arrange(time)

```




```{r}
hospital%>%
  get_mon_arrivals%>%
  mutate(waiting_time = end_time - start_time - activity_time)
```


#resource usage

#distribution of queue(nurse & room)


```{r}

plot(get_mon_resources(hospital), "usage", c("triage nurse", "room"), steps = TRUE)

```





#resources??? activity during the simulation.

```{r}
plot(resources, metric = "usage", c("triage nurse", "room"), items = "server")

```

#patient waiting time distribution

```{r}
arrivals <- get_mon_arrivals(hospital, per_resource = T)
summary(arrivals)
# Let's see the distributions
waiting_time <- arrivals$end_time - arrivals$start_time - arrivals$activity_time

```


#waiting_time 

```{r}
summary(waiting_time)

```

#distribution of waiting_time

```{r}
hist(waiting_time, freq = FALSE)
lines(density(waiting_time))
```


try to fit distribution of waiting_time


```{r}
library(fitdistrplus)
descdist(waiting_time, discrete = FALSE)

```

attempt to fit different distributions.
For example: normal distribution


normal_dist <- fitdist(waiting_time, "norm")
plot(normal_dist)





Next we can have a look at the evolution of the arrivals??? flow time during the simulation.
In the plot below, each individual line represents a replication. 
A smooth line is drawn over them.
All arrivals that didn’t finish their entire trajectory are excluded from the plot.

only one replication:

```{r}
arrivals <- get_mon_arrivals(hospital)
plot(arrivals, metric = "flow_time")

```

Similarly one can have a look at the evolution of the activity times 
with metric = "activity_time" and waiting times with metric = "waiting_time".

```{r}
arrivals <- get_mon_arrivals(hospital)
plot(arrivals, metric = "waiting_time")

```

#replications
waiting_time for each replication
calculate the mean 

# with stationary distribution arrivals
hospitalsims<-lapply(1:numReplications, function(i){
  simmer("hospital") %>%
  add_resource("triage nurse",numNurse) %>%
  add_resource("room", numRoom)%>%
  add_generator("patient", patient, function() {rexp(1, 1/7.28)})%>% 
    run(until = 1440)
})



#with Non stationary distribution arrivals
```{r}
hospitalsims<-lapply(1:numReplications, function(i){
  simmer("hospital") %>%
  add_resource("triage nurse",numNurse) %>%
  add_resource("room", numRoom)%>%
  #schedule(c(6, 12, 18, 24), c(3, 5, 2, 1), period=24)
  
    
  add_generator("patientESI1-a", patient1, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI1-b", patient1, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI1-c", patient1, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI1-d", patient1, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI2-a", patient1, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI2-b", patient1, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI2-c", patient1, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI2-d", patient1, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI3-a", patient1, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI3-b", patient1, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI3-c", patient1, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI3-d", patient1, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI4-a", patient1, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI4-b", patient1, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI4-c", patient1, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI4-d", patient1, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI5-a", patient1, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI5-b", patient1, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI5-c", patient1, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI5-d", patient1, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI0-a", patient1, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI0-b", patient1, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI0-c", patient1, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI0-d", patient1, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI1-e", patient2, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI1-f", patient2, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI1-g", patient2, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI1-h", patient2, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI2-e", patient2, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI2-f", patient2, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI2-g", patient2, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI2-h", patient2, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI3-e", patient2, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI3-f", patient2, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI3-g", patient2, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI3-h", patient2, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI4-e", patient2, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI4-f", patient2, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI4-g", patient2, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI4-h", patient2, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI5-e", patient2, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI5-f", patient2, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI5-g", patient2, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI5-h", patient2, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
  add_generator("patientESI0-e", patient2, from_to(0, 6*60, function() rexp(1, rate = 0.02)))%>%
  add_generator("patientESI0-f", patient2, from_to(6*60, 12*60, function() rexp(1, rate = 0.026)) )%>%
  add_generator("patientESI0-g", patient2, from_to(12*60, 18*60, function() rexp(1, rate = 0.029)) )%>%
  add_generator("patientESI0-h", patient2, from_to(18*60, 24*60, function() rexp(1, rate = 0.012)) )%>%
   
    
    
    
#generate the arrivals according to time
#can using from_to to change the parameters

 run(until = 1440)
})
```


```{r}
plot(get_mon_resources(hospitalsims))

```

#resources monitor
```{r}
resourcesmonitor <- lapply(1:numReplications, function(i){
  result<-hospitalsims[[i]]%>%
    get_mon_resources()
})


```


# arrival monitor
```{r}
arrivalmonitor<-lapply (1:numReplications, function(i){
  result<-hospitalsims[[i]]%>%
    get_mon_arrivals()%>%
    dplyr::mutate(service_start_time = end_time - activity_time) %>%
    mutate(wait_time = service_start_time - start_time)
})

```


#patient monitor
```{r}
patientmonitor <- lapply(1:numReplications, function(i){
        result <- hospitalsims[[i]] %>% 
          get_mon_arrivals %>%
          mutate(flow_time = end_time - start_time)
})


```


```{r}
get_mean_flow <- function(entitymonitor){
  mean(entitymonitor$flow_time)
}
```


```{r}
avg_patient_flow <- sapply(patientmonitor, get_mean_flow)
avg_patient_flowtime <- data.frame(run = 1:numReplications,
                                 avg_flow_time = avg_patient_flow)
avg_patient_flowtime
```


#get mean for waiting_time

```{r}

res <- get_mon_arrivals(hospitalsims, per_resource = TRUE) %>%
  dplyr::select(name, resource) %>%
  dplyr::filter(resource %in% c("triage nurse", "room"))


arr <- get_mon_arrivals(hospitalsims) %>%
  dplyr::mutate(waiting_time = end_time - (start_time + activity_time), 
                generator = regmatches(name,regexpr("patient[[:alpha:]]", name))
                ) %>%
  dplyr::left_join(res) %>%
  dplyr::group_by(generator, resource) %>%
  summarise(avgwait = mean(waiting_time, na.rm=TRUE))
```

```{r}
arr2 <- get_mon_arrivals(hospitalsims, per_resource = TRUE) %>%
  dplyr::mutate(waiting_time = end_time - (start_time + activity_time), 
                generator = regmatches(name, regexpr("patient[[:alpha:]]", name))) %>%
  dplyr::left_join(res)
```



```{r}
waitstatistics <- arr2 %>% group_by(resource) %>%
  dplyr::summarise(meanwait = mean(waiting_time, na.rm=TRUE),
                   sdwait = sd(waiting_time, na.rm=TRUE))
waitstatistics
```




#server utilization
```{r}
serverutilization <- function(resource){
  resource <- resource %>% arrange(time)
  lenreg <- nrow(resource)
  timein  <- c()
  for(i in 1:(lenreg-1)){
    timeinstate = resource$time[i+1] - resource$time[i]
    timein <- append(timein, timeinstate)
  }
  timein <- append(timein, 0)
  resource$timein <- timein
  serverutilization <- sum(resource$server * resource$timein) /(max(resource$time) - min(resource$time))
  serverutilization
}
```


#triage nurse monitor
```{r}
triagenursemonitor<-resourcesmonitor[[1]]%>%
  filter(resource == "triage nurse")%>%
  arrange(time)
triagenurseutilization<-serverutilization(triagenursemonitor)

triagenurseutilization
```



